---
title: "Static Data Exploration — Seoul Bike (Kaggle)"
format:
  html:
    toc: true
    toc-depth: 2
    theme: cosmo
    code-fold: true
execute:
  echo: true
  warning: false
  message: false
---
## Introduction

This report provides a static, reproducible exploration of the Seoul Bike Sharing Kaggle dataset. 
It mirrors the functionality of the Shiny app by showing categorical tables, numeric summaries, and 
a plot set (≥6 plots, ≥4 multivariate, ≥1 faceted, ≥1 novel).

**Dataset:** *Seoul Bike Sharing Demand Prediction* (Kaggle).  
**Unit of observation:** hour-level weather/operational records with `rented_bike_count`.

## Data Source & Variables (short)

Key variables used here:

- **Categorical:** `seasons`, `holiday`, `functioning_day`
- **Numeric:** `rented_bike_count`, `temperature_c`, `humidity`, `wind_speed_m_s`, `visibility_10m`

We convert `date` to Date and coerce the above categoricals to factors for tidy summaries/plots.

```{r}
library(readr); library(dplyr); library(tidyr); library(ggplot2); library(janitor)
library(forcats); library(stringr); library(purrr); library(scales)

# Read with UTF-8 locale to avoid "invalid UTF-8" warnings

df <- readr::read_csv(
  "data/SeoulBikeData.csv",
  show_col_types = FALSE,
  locale = readr::locale(encoding = "UTF-8")
) |> janitor::clean_names()

df <- df %>%
  mutate(
    date = suppressWarnings(as.Date(date, format = "%d/%m/%Y")),
    seasons = as.factor(seasons),
    holiday = as.factor(holiday),
    functioning_day = as.factor(functioning_day)
  )
```

## Structure

```{r}
glimpse(df)
summary(df)
```

## Categorical Summaries

### One-way table
We show counts by a single categorical variable (default: `seasons`).
Cross-tabulation of seasons × functioning_day, with row/column totals.
```{r}
one_way_var <- "seasons"
df %>%
  count(.data[[one_way_var]]) %>%
  arrange(desc(n)) %>%
  janitor::adorn_totals(name = "Total")

df %>%
count(seasons, functioning_day) %>%
tidyr::pivot_wider(names_from = functioning_day,
values_from = n, values_fill = 0) %>%
janitor::adorn_totals(where = c("row","col"))


```


